Chatbots and messaging platforms in the classroom: an analysis from the teacher's perspective

### 2023-01-09 edubots_surveys.Rnw
Update edubots_surveys.Rnw
Co-authored-by: Francisco Barranco <fbarranco@ugr.es>

\textcolor{black}{This study followed a quantitative approach towards addressing its research objective, by collecting rich data about the use of messaging applications and chatbots
by educators in both universities and colleges\footnote{In Spain, they are called simply high schools, and deliver
tertiary or associated degrees that do not have the consideration of university degrees}.
To achieve this goal two online surveys were designed and developed using Google Forms
in Spanish. They are comprised of six questions focusing on demographic data
(i.e., sector, gender, degrees, discipline, age, teaching experience)
followed by several multiple choice questions allowing the participants to choose multiple answers.}

In the {\bf first survey}, the multiple choice questions are
focused on the use of messaging apps in the teaching practice, the type of chatbot use cases that teachers consider
useful for their teaching and the impact of COVID on the teaching practice; please check them out to the full extent in the Appendix.\footnote{This {\bf first survey} was made in two separate Google forms, one for
university and the other for teachers in tertiary education institutions, outside the university system (vocational training, including 1- and 2-year degrees). Questions and
responses were the same, except for the type of tertiary degrees that
were considered.}

\textcolor{black}{This survey was piloted by the authors of the paper and their colleagues before using it in the
study.  All the comments collected from this trial were incorporated in the survey.
We mainly used mailing lists, as well as Telegram groups,
to post the survey link to reach out to more educators. The university form was sent to Spanish
(mainly Andaluc\'{\i}a and Galicia) universities, and also  universities in Costa Rica and
Mexico. The tertiary (non-university/college) teachers were based mainly in
Andaluc\'{\i}a. Dissemination and answers took place in the first quarter of
2021, during the Covid-19 pandemic and while, at least in Spain, many universities had
mandatory virtual teaching.}

A total of 282 teachers responded to the survey: 193 teaching at the University
(68.4\%) and 89 at other tertiary education institutions (31.6\%). \textcolor{black}{From those, 179 indicated their gender as male (63.5\%), 98 as female (34.8\%) and 5 teachers preferred not to indicate their gender (1.8\%). In terms of age, the majority of the participants (n=111) in this survey was 45-55 years old (39.4\%), while 91 teachers were 35-45 (32.3\%), 46 were 25-35 (16.3\%) and 34 teachers were older than 55 (12.1\%). Last, 84 teachers had a teaching experience of 16-25 years (29.8\%), 75 teachers 6-15 years of experience (26.6\%), 69 teachers had 0-5 years of experience (24.5\%) and finally, 54 teachers had more than 25 years of teaching experience (19.1\%).}

@@

The most favorable use case for chatbots in
Higher Education and vocational training is their use for the facilitation of an
agenda formation (171 positive response, 60.6\%), followed by the FAQs use case
(148 positive responses, 52.5\%) and the sharing class material use case (136;
48.2\%). A chi-square test of independence was performed to examine the relation
among participants’ preferences for particular chatbot use cases. Out of the 171 teachers who consider useful the use of
chatbots for agenda preparation in the class, 103 also consider useful chatbots’
use for FAQs. The relation between agenda and the FAQs use case was significant, $X^{2}_{(1, N = 282)} =
10.467$, $p =0.001$. The frequencies cross tabulated are given in Table
\ref{tab:agenda*faq}.

@@

Our initial intention in the design of these surveys was to probe the opinions
of tertiary education teachers in the introduction of chatbot technologies in
class. We wanted to find the answer to four research questions, of which the
first was to check whether teachers already used some kind of messaging
application, and the one they preferred. Depending on the answer, different
technologies could be (or not) available for the introduction of chatbots. In
this case, the design of the survey introduced some ambiguity on what we really
consider ``messaging technology'', that was eventually resolved by the answers.
Apparently, anything that sends messages (even if they are not instant, even if
they do not have a dedicated application) was the concept in the minds of the
surveyed. In this case, we would like
to add a clarification. While we had in mind, as indicated by the possible answers,
instant messaging applications when we elaborated this question,
we included these institution-provided applications mainly for completion. But
it should be noted that, in most cases, they are not {\em instant} messaging
apps in the way Telegram or Facebook Messenger are; in most cases, they are
messaging facilities provided by learning management systems such as (usually)
Moodle. So the answer to this question must be understood in two different ways:
their preference for the {\em provider} of the technology (the university itself, any
company) as well as the {\em type} of technology (instant, app-based messaging
systems vs. web-based messaging systems integrated in another learning management system).

This inclusion of university-provided messaging applications in
the survey might explain why there are so few teachers, just 6.7\%, that answered they did not
use any messaging application. In general, universities provide official channels
to communicate with students official academic information such as the exam schedule
or the grades. The statistical mode is to use that kind of messaging. Some other
teachers might not use it, but use some extraofficial or opt-in channels such as the others included in the survey; in the end, between these official and extra-official channels, most people cover their needs to communicate with students. This matches
our subjective experience, in general.

Additionally, the question on COVID reflected the big challenge that represents
changing habits, even in the case of a major crisis. Very few responders actually
changed their habits, although comparing women vs. men as well as those teaching social
sciences show some significant differences. We do not have any kind of explanation for
this behavior, other than technologies are perceived, and adopted, in a
different way depending on background and, apparently, gender.

One of the most interesting questions arising from the analysis of the
results, and that should require further study, is: Are teachers really aware
of the possibilities of chatbots in the classroom? The survey
only listed a few functionalities for the chatbots, and just a few subjects selected
the option "Others". This may suggest that most teachers did not really understand
at this point the actual potential of chatbots, since the responses listed were by no means exhaustive. If the responders were more acquainted with the possibilities of this technology, this response would possibly have shown up many more times. Another way to look at it is that, unless you provide the specific instructions on how to use a
tool, the teaching staff is already too busy or burnt out to think about innovation.

We can speculate that teachers' opinions of chatbots are influenced by negative
experiences with customer support chatbots, which is what most people are likely to have
experienced. Further study may be needed to confirm this, but it is important to 
keep in mind when introducing chatbot technologies in the classroom.
To potentially change this existing perception, it will be necessary to make the chatbot as simple and satisfying to use as possible.

Regarding gender grouping, both genders agree in their uses, so it seems not
relevant at all. However, when we look at the plot containing the sector differentiation,
it is interesting to note that university teachers are more likely to use chatbots 
for grades and FAQs compared to other tertiary education teachers.
This seems reasonable as the frequency of seeing the students
is lower at the University so, an automatic tool for grade notification for
continuous evaluation would be more helpful.
Furthermore, the high student-to-teacher
ratios that university teachers often face could be another reason why
a higher number of them are willing to use chatbots for FAQs, as this could
reduce misunderstandings and email overload from students.

The results, when grouping by teaching experience, confirm that
experienced teachers are just as open to change and willing to embrace new
technological challenges as younger teachers. 
From this concrete plot it is interesting as well to observe the fact that very few
young teachers are willing to use chatbots for other uses. This
reinforces the hypothesis thrown at the beginning of this section that younger
teachers have less experience to consider new ways of using chatbots in their daily work.

According to the responses presented above, teachers’ responses are evenly
split between positive and negative in relation to the media features:
analytics, connectivity, and hidden phone number. This means that none of these features is a
deciding factor in their choice of technology. However, for features such as familiarity,
horizontality, sustainability, and officiality the majority provided a positive
response, designating their preference to these features. On the other hand, the majority of teachers
do not seem to see the media features of official certification, pluggability, unidirectionality, 
and synchrony as valuable in messaging apps.

From a qualitative point of view, these results show consensus in a few key aspects:
\begin{itemize}
 \item Teachers want to have support from their institutions. This
   is probably due to the strict European data regulations on data protection,
   but also to avoid the overhead required to sign up new students every year.
   In general, additional support with specific formation and access to IT help desk for specific features
   enhance the possibility of a successful adoption that improves the learning outcomes. This is supported by the result of the two
   surveys: they prefer whatever "messaging" application is provided by their university, and
   they the "officiality" of the platform, as shown in
   Table \ref{tab:interaction}.
 \item Teachers need sustainability. The main reason is that implementing
   changes is expensive in terms of effort and prone to errors in the first
   years. Therefore, it is logical that if
   chatbots or messaging platforms are to be introduced, teachers want this change
   to be as permanent as possible.
 \item Keep it Simple. Media interaction requirements are simple and only 2.4\% of teachers are requesting more features. This could be a sign of technological burnt out produced by the previous pandemic courses, when the use of computers significantly increased.
\end{itemize}

The analysis of the questions related to RQ4 seems to indicate that teachers
prefer to share an interaction space with their students. As
shown in Tables \ref{tab:kindinteractions} and \ref{tab:interaction}, the majority of responders
indicated that they {\em do not} want unidirectional communication, and that they prefer
{\em horizontal} interaction. This point is reinforced by results on the kind of
chatbots they preferred: most of them indicated FAQ bots, which might indicate a
need to offload part of the burden of answering every single question posed by
their students. In a horizontal setting, other students would answer if the teacher
does not do it immediately. This might also be the reason why the possibility of
using tools from which analytics can be extracted is also valued by teachers:
in that way, the student activity in the chat tool, answering questions, or their regularity can all be
assessed as part of the process of achieving learning objectives.

% Only some COVID-response relationship.
In our survey, we could not find an answer about the right moment to introducing chatbot technologies. The response to questions regarding to change of
behavior during and after the first stay-at-home stage of the pandemic do not suggest that a major (or minor, such as new higher education laws or the introduction of new degrees) crisis could be an opportunity to introduce new technologies. The reason being that it does not
bring major changes in attitude 
%(as a minor "crisis" we could mention any institutional event that brings about methodological changes, such as new higher education laws\footnote{We seem to have one of these, at least in Spain, every 10 years or so.}), 
%or simply the introduction of new degrees. 
Although changes driven by external forces do offer the chance of piloting new technologies, they do not
seem to bring internal changes in teachers' attitudes (which, after all, is
bound to be the same). In absence of a clear answer in this direction, the right
moment to introduce new technologies is always when the IT and managing staff is
ready to support it (since, as we have seen before, "official" support is one of
the factors that is most valued by teachers).

@@

The key findings of this study shed light on educators’ preferences for using
messaging applications supported by their institutions. Technology adoption
literature often focuses on users’ perceptions of the technology usefulness and
ease of use as important factors for its successful adoption and
utilization. Nevertheless, in higher education, institutions play a role in
integrating these tools into their educational systems, which can improve the uptake of
these applications and shape the social and educational experiences of 
students. To achieve this successful integration, institutions should ensure that
these messaging applications GDPR- (General Data Protection Regulation) compliant to protect students’ data and 
provide IT support to all stakeholders who use the applications.

Comparing these results with those obtained from student surveys
in \cite{MoraChatbots2021:anon}, we have better understood the differences between teachers’
and students' points of view and intentions when using the
messaging applications in higher education.  Teachers tend to adopt technologies
supported by their institutions, which could be motivated by a desire to ensure that their
universities can oversee their efforts in supporting their students during the
learning process.  Another reason could be their familiarity with the technology
provided by their institutions. Meanwhile, students use 
non-institution messaging applications to create informal discussion groups with
their peers.  It is worth noting that peer support and collaboration are
inseparable from learning \cite{timmis2012constant} and correlates positively
with higher retention in higher education \cite{o2014front}.  Therefore, both
perspectives are complementary and play different roles in promoting the
learning process.

As with many other methodologies for assessing technology acceptance, the survey results suggest
that the introduction of simple, and institution-supported instant messaging
and chatbot technologies would increase the perceived usefulness (which is one
of the key metrics in technology acceptance models). Given that most institutions
already have some {\em virtual campus} or learning management system, adding
some easy automation or connection to personal instant messaging tools could
help introduce these new technologies to the learning community.

Based on the discussion of this study, we propose a potential process for introducing
technology that would need to be piloted in order to determine its value and its
relationship to improved learning outcomes and higher teacher satisfaction.

Initially, institutions should introduce messaging automation tools that help with bureaucratic or repetitive tasks, such as answering frequently asked questions or providing information on class or assignment deadlines. 
These chatbots will pave the way for more complex chatbots that will
have a greater impact on learning outcomes, such as chatbots that help students
integrate in class or identify and address learning problems
in individual students or groups of them. These chatbots should also be accompanied by
analytics on student interaction and possibly some natural language
processing (in vernacular language) to help assess the general mood of
the class and how different materials or external factors affect it.

>>>>

This study has employed a quantitative approach in order to address its research objective. A good amount of data was collected on the usage of messaging applications and chatbots by educators in universities and colleges.\footnote{In Spain, higher education institutions generally known elsewhere as colleges, are called simply high schools, and deliver
tertiary or associate degrees that do not have the consideration of university degrees.}
To achieve this goal two online surveys were designed and developed using Google Forms
in Spanish. They are comprised of six questions focusing on demographic data
(i.e., sector, gender, degrees, discipline, age, teaching experience)
followed by several multiple choice questions allowing the participants to choose multiple answers.

In the {\bf first survey}, the multiple choice questions are
focused on the use of messaging apps in the teaching practice, the type of chatbot use cases that teachers would consider
useful for their teaching and the impact of COVID on the teaching practice; please check them out to the full extent in the Appendix.\footnote{This {\bf first survey} was made in two separate Google forms, one for
university and the other for teachers in tertiary education institutions, outside the university system (vocational training, including 1- and 2-year degrees). Questions and
responses were the same, except for the type of tertiary degrees that
were considered.}

This survey was piloted by the authors of the paper and their colleagues before using it in the
study.  All the comments collected from this trial were incorporated in the survey.
We primarily used mailing lists and Telegram groups to disseminate the survey link so that it would reach a larger number of educators. The form addressed at university teachers was sent to Spanish
(mainly Andaluc\'{\i}a and Galicia) universities, and also  universities in Costa Rica and
Mexico. The tertiary (non-university/college) teachers who received the form were based mainly in
Andaluc\'{\i}a. Dissemination and answers took place in the first quarter of
2021, during the Covid-19 pandemic and while, at least in Spain, many universities had
mandatory virtual teaching.

A total of 282 teachers responded to the survey: 193 teaching at the University
(68.4\%) and 89 at other tertiary education institutions (31.6\%). From those, 179 indicated their gender as male (63.5\%), 98 as female (34.8\%) and 5 teachers preferred not to indicate their gender (1.8\%). In terms of age, the majority of the participants (n=111) in this survey was 45-55 years old (39.4\%), while 91 teachers were 35-45 (32.3\%), 46 were 25-35 (16.3\%) and 34 teachers were older than 55 (12.1\%). Last, 84 teachers had a teaching experience of 16-25 years (29.8\%), 75 teachers 6-15 years of experience (26.6\%), 69 teachers had 0-5 years of experience (24.5\%) and finally, 54 teachers had more than 25 years of teaching experience (19.1\%).

@@

The most beneficial use case for chatbots in higher education and vocational training is their ability to facilitate agenda formation and communication (171 positive responses, 60.6\%), followed by the FAQs use case
(148 positive responses, 52.5\%) and the sharing class material use case (136;
48.2\%). A chi-square test of independence was performed to examine the relation
among participants’ preferences for particular chatbot use cases. Out of the 171 teachers who consider useful the use of
chatbots for agenda preparation in the class, 103 also consider useful chatbots’
use for FAQs. The relation between agenda and the FAQs use case was significant, $X^{2}_{(1, N = 282)} =
10.467$, $p =0.001$. The frequencies cross tabulated are given in Table
\ref{tab:agenda*faq}.

@@

Our initial intention in the design of these surveys was to probe the opinions
of tertiary education teachers in the introduction of chatbot technologies in
class. We wanted to find the answer to four research questions, of which the
first was to check whether teachers already used some kind of messaging
application, and the one they preferred. Depending on the answer, different
technologies could be (or not) available for the introduction of chatbots. In
this case, the design of the survey introduced some ambiguity on what we really
consider ``messaging technology'', that was eventually resolved by the answers;
apparently, anything that sends messages (even if they are not instant, even if
they do not have a dedicated application) was the concept in the mind of the
surveyed.  \textcolor{black}{In this case, we would like
to add a clarification. While we had in mind, as indicated by the possible answers,
instant messaging applications when we elaborated this question,
we included these institution-provided applications mainly for completion. But
it should be noted that, in most cases, they are not {\em instant} messaging
apps in the way Telegram or Facebook Messenger are; in most cases they are
messaging facilities provided by learning management systems such as (usually)
Moodle. So the answer to this question must be understood in two different ways:
their preference for the {\em provider} of the technology (the university itself, any
company) as well as the {\em type} of technology (instant, app based messaging
systems vs. web-based messaging systems integrated in another learning management system).}

This inclusion of university-provided messaging applications in
the survey might explain why there are so few, just 6.7\%, that answer they do not
use any messaging application. In general, universities provide official channels
to communicate with students official academic information such as the exam schedule
or the grades. The statistical mode is to use that kind of messaging. Some other
people might not use it, but use some extraofficial or opt-in channels such as the other included in the survey; at the end, between these official and extraofficial channels most people cover their needs to communicate with students. This matches
our subjective experience, in general.

Additionally, the question on COVID reflected the big challenge that represents
changing habits, even in the case of a major crisis. Very few responders actually
changed their habits, although comparing women vs. men as well as those teaching social
sciences show some significant difference. We do not have any kind of explanation for
this behavior, other than technologies are perceived, and adopted, in a
different way depending on background and, apparently, gender.

One of the most interesting questions that can arise from the analysis of the
results, and that should require further study, is: Are teachers really aware
of the possibilities of chatbots in the classroom? The survey
only listed a few functionalities for the chatbots and just a few subjects selected
the option "Others". This may suggest  that most teachers did not really understand
at this point the actual potential of chatbots, since the responses listed were by no means exhaustive, and if the responders were more acquainted with the possibilities of the technology, possibly this response would have shown up many more times. Another way to look about it is
that, unless you provide the specific instructions on how to use a
tool, the teaching staff is already too busy or burn out to think about
innovation.

We can speculate that their opinions towards chatbots are informed by negative
experiences with customer support chatbots, which is what most people will have
experienced. This will probably require further study, but also will need to be
taken into account when introducing chatbot technologies in the classroom:
it will have to be as simple, and as satisfactory, as possible,
to (possibly) change this existing perception.

Regarding gender grouping, both genders agree in their uses, so it seems not
relevant at all. When looking at the plot containing the sector differentiation
it is interesting to observe how university teachers prefer in a higher
proportion to use them for grades and FAQs in comparison with other tertiary education teachers.
This seems reasonable as the frequency of seeing the students
is lower at the University so, an automatic tool for grade notification for
continuous evaluation seems more helpful.
Furthermore, the high student/teacher
ratios that university teachers have to deal with could be another reason to
have a higher number of them willing to use chatbots for FAQs. This could
ameliorate the number of misunderstandings and email overload that so many
students might generate.

The results, when grouping by experience, confirm that
experienced teachers are not afraid of changes and are willing to accept new
technological challenges; at least, in the same way that younger teachers
are. From this concrete plot it is interesting as well to observe the fact that very few
young/inexperience teachers are willing to use chatbots for other uses. This
reinforces the hypothesis thrown at the beginning of this section that younger
teachers have less experience to consider new ways of using chatbots in their daily work.

As it appears in the responses presented above, teachers’ responses are equally
distributed between positive and negative in relation to the media features:
analytics, connectivity, and hidden phone number; that is, none of them is a
prevailing factor on choosing one technology over another. However, for the features: familiarity,
horizontally, sustainability and officiality the majority provided a positive
response, designating their preference to these features. On the other hand, the
majority of teachers seems to consider not valuable the presence of official
formation, pluggability, unidirectionality and synchrony in the interaction
media features of messaging apps.

From a qualitative point of view, these results show consensus in a few key aspects:
\begin{itemize}
 \item Teachers want to have support from their corresponding institution. This
   is probably due to the strict European data regulations \textcolor{black}{on data protection},
   but also to avoid the overhead required to \textcolor{black}{sign up new students every year.
   In general support from specific formation and access to IT help desk for specific features
   enhances the possibility of adoption by the teachers or the possibility of a successful
   adoption that improves the learning outcomes. This is supported by the result of the two
   surveys: they prefer whatever "messaging" application is provided by the university, and
   also they value sustainability as well as "officiality" of the platform, as shown in
   Table \ref{tab:interaction}}.
 \item Teachers need sustainability. The main reason is that implementing
   changes is expensive in terms of effort and prone to errors in the first
   years of implementing new procedures, therefore, it is logical that if
   chatbots or messaging platforms are to be introduced, they want this change
   to be as permanent as possible.
 \item Keep it Simple. Media interaction requirements are simple and only a 2.4\% is requesting more features. This could be a sign of technological burn out produced by the previous and current courses where, due to the pandemic, the use of computer has increased significantly.
\end{itemize}

An analysis of survey questions related to RQ4 seems to indicate that teachers
prefer to share an interaction space with their student, as
shown in Tables \ref{tab:kindinteractions} and \ref{tab:interaction}, where a majority of responders
indicate they {\em don't} want unidirectional communication; a preference of
{\em horizontal} interaction, but also the answer to the kind of
chatbots preferred that features FAQ bots prominently, which might indicate a
need to delegate or offload part
of the burden of answering, and attending, every single question posed by
students. In an horizontal setting, other students can answer if the teacher
does not do it immediately. This might also be the reason why the possibility of
using tools from which analytics can be extracted is also valued by teachers:
that way, activity in the chat tool, answering questions, regularity, can all be
observed and valued as part of the process of assessing achievement of learning
objectives by students.

% Only some COVID-response relationship.
In our survey, we could not find an answer about the right moment for introducing chatbot technologies. Response to questions related to change of
behavior during and after the first stay at home, stage of the pandemic do not suggest that a major (or minor)
crisis could be an opportunity to introduce new technologies, since it does not
bring major changes in attitude (as a minor "crisis" we could mention any institutional event that brings about methodological changes, such as new higher education laws\footnote{We seem to have one of these, at least in
  Spain, every 10 years or so.}), or simply introduction of new degrees. Although
changes driven by external forces do offer the chance of piloting new technologies, they do not
seem to bring internal changes in attitudes in the teaching staff (which, after all, is
bound to be the same). In absence of a clear answer in this direction, the right
moment to introduce new technologies is always when the IT and managing staff is
ready to support it (since, as we have seen before, "official" support is one of
the factors that is most valued by teachers).

@@

The key findings of this study shed light on educators’ preference to use
messaging applications that their institutions support. Technology adoption
literature often \textcolor{black}{focuses} on users’ perceptions of the technology’s usefulness and
ease of use as important prerequisites for successful adoption and
utilization. Nevertheless, in higher education, institutions’ role in
integrating these tools to their educational systems can improve the uptake of
these applications and shape the social and educational experiences of their
students. \textcolor{black}{To accomplish this fruitful integration}, institutions have to ensure that
these messaging applications not only are GDPR (General Data Protection Regulation) compliant to keep students’ data
secure, but also should provide IT support to all stakeholders who use the
applications.

Comparing these results with those obtained in student surveys
in \cite{MoraChatbots2021:anon}, we have understood better the differences between the teachers’
points of view and/or intentions and that of their students when using the
messaging applications in higher education.  Teachers tend to adopt technologies
supported by their institutions.  The fact they do so could be caused by a desire to ensure that their
universities can oversee their efforts in supporting their students during the
learning process.  Another reason could be their familiarity with the technology
provided by their institutions. On the other hand, students use the
non-institution messaging applications to form informal discussion groups with
their peers.  It is worth noting that peer support and collaboration is
inseparable from learning \cite{timmis2012constant} and correlates positively
with higher retention in higher education \cite{o2014front}.  Therefore, both
perspectives are complementary and play different roles in promoting the
learning process.

As in many other methodologies that try to assess technology acceptance, answers to the survey suggest
that the introduction of simple, and institution-supported, instant messaging
and chatbot technologies would increase the perceived usefulness (which is one
of the key metrics in technology acceptance models). Since most institutions
already have some {\em virtual campus} or learning management system, adding
some easy automation, or connection to personal instant messaging tools, could
really help onboard the learning community on these new technologies.

What we are going to propose next is a possible process of
technology introduction that is compatible with the conclusions of this study,
but that would have to be piloted in order to check its value, and its
relationship with better learning outcomes, as well as higher teacher
satisfaction.
Once that initial introduction of institutional messaging
automation tools is done, teachers (and students as well) will
probably prefer the kind of bots that alleviate bureaucratic or repetitive
tasks, such as answering frequent questions or answers on class or assignment
deadlines, as indicated by their answers to the respective
questions. These will help the introduction of more complex chatbots that will
affect more directly learning outcomes, such as chatbots that help students
integrate in the class, or are able to identify (and address) learning problems
in students or in groups of them. These should also be accompanied by
analytics on student interaction, as well as possibly some natural language
processing (in vernacular language) that will help assess the general mood of
the class, and how different material (or external factors) affect it.
